---
title: "Chocolate Chip Cookies"
execute:
  error: true
author: "Sowparnika"
output: html_document
---

## Reading In the Data

First, read in the CSV data of cookie ingredients.
Make sure that your end-result data has appropriate types for each column - these should match the types provided in the documentation in the README.md file.

```{r}
cookies <- read.csv("choc_chip_cookie_ingredients.csv")
cookies |> str()
```


```{python}
import pandas as pd
cookies = pd.read_csv("choc_chip_cookie_ingredients.csv")
cookies.info()
```


## Exploratory Data Analysis

Exploratory data analysis is the process of getting familiar with your dataset. To get started, [this blog post](https://www.mrdbourke.com/a-gentle-introduction-to-exploratory-data-analysis/) provides a nice checklist to get you thinking:

> 1.  What question(s) are you trying to solve (or prove wrong)?
> 2.  What kind of data do you have and how do you treat different types?
> 3.  What's missing from the data and how do you deal with it?
> 4.  Where are the outliers and why should you care about them?
> 5.  How can you add, change or remove features to get more out of your data?

### Generating Questions

Generate at least 5 questions you might explore using this database of cookie ingredients.

1. Which ingredients are rarely used?
2. Do certain ingredients commonly appear together?
3. Which ingredients appear most frequently? 
4. How many ingredients does the average recipe contain?
5. Which recipes use the largest number of different ingredients?


### Skimming the Data

One thing we often want to do during EDA is to examine the quality of the data - are there missing values? What quirks might exist in the dataset?

The `skimr` package in R, and the similar `skimpy` package in python (which has a much better name, in my opinion), can help provide visual summaries of the data. 

Install both packages, and read the package documentation ([R](https://cran.r-project.org/web/packages/skimr/vignettes/skimr.html), [Python](https://pypi.org/project/skimpy/)).

[Part 1] Use each package and generate summaries of your data that require the use of at least some non-default options in each package's `skim` function.


```{r}
library(skimr)
skim(cookies)

```

```{python}
from skimpy import skim
skim(cookies)
```

[Part 2] Write 1-2 sentences about what you can tell from each summary display you generate. Did you discover anything new about the data?

R-
The summaries show that nearly half of the Rating values are missing, which may affect analysis involving ratings. The Quantity variable varies widely, suggesting differences in ingredient amounts across recipes.

Python-
The dataset contains both numeric and string variables, with ingredient names and units showing consistent formatting but varied lengths.

### Generating Tables

Another useful technique for exploratory data analysis is to generate summary tables. 
You may want to use the `dplyr` package in R (`group_by` or `count` functions), as well as the `groupby` and `count` methods in Pandas. [Python example](https://sparkbyexamples.com/pandas/pandas-groupby-count-examples/), [R example](https://dplyr.tidyverse.org/reference/count.html)

[Part 1] Using R and Python, generate a table that shows what **proportion** of recipes contain each type of ingredient, for the most common 20 ingredients.
```{r}
library(dplyr)
total_recipes <- n_distinct(cookies$Recipe_Index)
ingredient_props <- cookies %>%
  distinct(Recipe_Index, Ingredient) %>%      # remove duplicate ingredient entries per recipe
  count(Ingredient, sort = TRUE) %>%    # count recipes per ingredient
  mutate(proportion = n / total_recipes) %>%
  slice_head(n = 20)

ingredient_props

```

```{python}
import pandas as pd

total_recipes = cookies["Recipe_Index"].nunique()

ingredient_props = (
    cookies[["Recipe_Index", "Ingredient"]]
    .drop_duplicates()
    .groupby("Ingredient")
    .size()
    .reset_index(name="count")
)

ingredient_props["proportion"] = ingredient_props["count"] / total_recipes

ingredient_props = ingredient_props.sort_values(
    by="count", ascending=False
).head(20)

ingredient_props


```

[Part 2] Print out a character string that lists all of the ingredients that do not appear in at least 20 recipes.
```{r}
library(dplyr)

# Count number of recipes per ingredient
ingredient_counts <- cookies %>%
  distinct(Recipe_Index, Ingredient) %>%
  count(Ingredient)

# Find ingredients appearing in fewer than 20 recipes
rare_ingredients <- ingredient_counts %>%
  filter(n < 20) %>%
  pull(Ingredient)

# Print as a single character string
paste(rare_ingredients, collapse = ", ")

```
```{python}
# Count recipes per ingredient
ingredient_counts = (
    cookies[["Recipe_Index", "Ingredient"]]
    .drop_duplicates()
    .groupby("Ingredient")
    .size()
    .reset_index(name="count")
)

# Filter ingredients with fewer than 20 recipes
rare_ingredients = ingredient_counts[
    ingredient_counts["count"] < 20
]["Ingredient"]

# Print as a single string
print(", ".join(rare_ingredients))

```

### Visualization

Using whatever plotting system you are comfortable with in R or python, see if you can create a couple of useful exploratory data visualizations which address one of the questions you wrote above - or another question which you've come up with as you've worked on this assignment.

[Part 1] Create at least one plot (it doesn't have to be pretty) that showcases an interesting facet of the data.
```{r}
library(dplyr)
library(ggplot2)

ingredients_per_recipe <- cookies %>%
  distinct(Recipe_Index, Ingredient) %>%
  count(Recipe_Index)

ggplot(ingredients_per_recipe, aes(x = n)) +
  geom_histogram(binwidth = 1) +
  labs(title = "Number of Ingredients per Recipe",
       x = "Number of Ingredients",
       y = "Number of Recipes") +
  theme_minimal()

```

[Part 2] Write 2-3 sentences about what you can learn from that plot and what directions you might want to investigate from here.

Most recipes have around 9–10 ingredients, so that seems to be the typical number for chocolate chip cookies. Only a few recipes use very few or very many ingredients. This suggests that most recipes follow a similar basic structure, and it might be interesting to see which ingredients are always included and which ones vary.
